---
title: "Bridge Math Final"
author: "Ahsanul Choudhury"
date: "August 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=TRUE}
#install.packages("hflights")

#loading requird packages
require (hflights)
require (ggplot2)
head (hflights)
```

#Define Variables
```{r variables, include=TRUE}
#converting all the na values to zero for consistancy
hflights$ArrDelay[ is.na( hflights$ArrDelay ) ] <- 0
hflights$DepDelay[ is.na( hflights$DepDelay ) ] <- 0

X <- hflights$ArrDelay
Y <- hflights$DepDelay
```

#Generate summary level descriptive statistics
```{r summary, include=TRUE}
summary (X)

summary (Y)


Xtotal <- length(X) - 3622
Xtotal

Ytotal <- length(Y) - 2905
Ytotal
```


#Density plot for X and Y
```{r histA, include=TRUE}
ggplot(hflights, aes(X)) +
  geom_density()

ggplot(hflights, aes(Y)) +
  geom_density()
```

#'x' as 3d quartile of 'X' and 'y' as 2d quartile 'Y'
```{r quartile, include=TRUE}
x <- quantile(X, .75)
x

y <- quantile(Y, .5)
y
```


#Table of counts
```{r table, include=TRUE}
a <- nrow(subset(hflights, X <= x & Y <= y));
a;

b <- nrow(subset(hflights, X <= x & Y > y));
b;

c <- nrow(subset(hflights, X > x & Y <= y));
c;

d <- nrow(subset(hflights, X > x & Y > y));
d;


values <- matrix(c(a,b,(a+b),c,d,(c+d),(a+c),(b+d),(a+b+c+d)),ncol=3, nrow=3,byrow=TRUE)
colnames(values) <- c("<=2d quartile",">2d quartile","Total")
rownames(values) <- c('<=3d quartile', '>3d quartile','Total')
values.table <- as.table(values)
values.table
```

# a. P(X>x | Y>y)         b.  P(X>x, Y>y)         c.  P(X<x | Y>y)
```{r sol1, include=TRUE}
# a. P(X>x | Y>y)
48548/109996

# b.  P(X>x, Y>y)
48548/227496

#c.  P(X<x | Y>y)
(nrow(subset(hflights, X < x & Y > y)))/(nrow(subset(hflights, Y > y)))
```

#Does P(A | B) = P(A) * P(B)?
```{r sol2, include=TRUE}
#P(A | B) = P(X>x | Y>y)
48548/109996
#P(A)
A <- nrow(subset(hflights, X > x))
A

pofA <- A/nrow(hflights)
pofA

#P(B)
B <- nrow(subset(hflights, Y > y))
B


#P(A | B) is not equal to P(A) * P(B)
pofB <- B/nrow(hflights)
pofB
#P(A) * P(B)
p <- pofA * pofB
p
# P(A | B) = .4413 in not equal to P(A) * P(B) = .1163(X and Y variable are not independent)
```

```{r chi, include=TRUE}
#Chi Squere Test for association

chisq.test(X,Y)

#We can see form the Chi Square test the p value is smaller than significance level of 0.05, we should reject the hypothesis that X (Arrival delay) and Y Departure dealy are independent
```

#scatterplot of the two variables
```{r scatt, include=TRUE}
qplot(X, Y, main = 'Relationship between Arrival dealay and Departure dalay',
      xlab = 'Arrival Delay', ylab = 'Departure Delay')
#Scatter plot indicates strong positive relationship between the two variable
```

#95% confidence interval for the difference in the mean
```{r ttest, include=TRUE}
t.test(X,Y)
#95% confidence interval for the difference in mean -2.514811 to -2.171108. Values are negative because the mean of random variable X(6.981384) is smaller then mean of random variable Y(9.324344)
```

#correlation matrix for two of the quantitative variables 
```{r cmatrix, include=TRUE}
data <- cbind(X,Y)
matrix <- cor(data, use = "complete")
matrix
```

#Hypothesis test to see if correlation between these variables is 0 and provide a 99% confidence interval
```{r hypotest, include=TRUE}
cor.test(X,Y,conf.level=0.99)
# 99% confidence interval is 0.924667 to 0.9262172. Correlation between X and Y is not zero, therefore the variables are not independent. Correlation coeffieient of 0.9254458 points to strong dependance.
```


#Invert the Correlation Matrix to Create the 'Precision Matrix'
```{r invert, include=TRUE}
pmatrix <- solve(matrix)
pmatrix
```

#Multiply the Correlation Matrix by the Precision Matrix
```{r multiply1, include=TRUE}
matrix%*%pmatrix
```
#Multiply the Precision Matrix by the Correlation Matrix
```{r multiply2, include=TRUE}
pmatrix%*%matrix
#The results of both multiplication are inverse of each other
```

#shift X so that the minimum value is above zero
```{r shift, include=TRUE}
Xshift  <- X+ abs(min(X)) + 1
summary (Xshift)
```

#Load the MASS package
```{r mass, include=TRUE}
require(MASS)
library(MASS)
```

#run fitdistr to fit an exponential probability density function
```{r fitdistr, include=TRUE}
Xexpo <- fitdistr(Xshift, "exponential")
Xexpo
```

#Optimal value of lambda for the distribution
```{r lambda, include=TRUE}
1 / Xexpo$estimate
```

# 1000 samples from the exponential distribution
```{r sample, include=TRUE}
sampex <- rexp(1000, Xexpo$estimate)
summary(sampex)
```

#Comparison of histogram of expotential sample with a histogram of original variable
```{r histogram, include=TRUE}
qplot(sampex, geom="histogram", binwidth = 5, fill=I("blue"), col=I("black"))
qplot(X, geom="histogram", binwidth = 5, fill=I("blue"), col=I("black"))
#The expotential sample appears to be less right skwed then the original
```

# 5th Percentile
```{r fifth, include=TRUE}
qexp(.05, rate = Xexpo$estimate, lower.tail = TRUE, log.p = FALSE)
```

#95th Percentile
```{r 95th, include=TRUE}
qexp(.95, rate = Xexpo$estimate, lower.tail = TRUE, log.p = FALSE)
```


#95% confidence interval from the empirical data, assuming normality
```{r 95thconf, include=TRUE}
meanX <- mean(na.omit(X)) # get the mean for X
meanX
sdX <- sqrt(var(na.omit(X)))
sdX

qnorm(.025, mean = meanX, sd = sdX)

qnorm(.975, mean = meanX, sd = sdX)
#If X is normally distributed 95% of the sample mean will be between -52.75051 and 66.71328
```


#The empirical 5th percentile and 95th percentile of the data
```{r empirical, include=TRUE}
#.05 quantile 
quantile(X, .05, na.rm = TRUE)

#.95 quantile
quantile(X, .95, na.rm = TRUE)

#GThe empirical 5th percentile and 95th percentil of the implies 90% of the sample X from original data falls between -18 and 56
```