---
title: "R Final"
author: "Ahsanul Choudhury"
date: "August 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#birthwt(Risk Factors Associated with Low Infant Birth Weight)

#Description

#The birthwt data frame has 189 rows and 10 columns. The data were collected at Baystate Medical Center, Springfield, Mass during 1986.

#Source: https://vincentarelbundock.github.io/Rdatasets/datasets.html

#In this R project I will use the birthwt dataset to see if there is any raltionship between mother smoking during pregnancy and baby being under weight at birth.

#required packages
```{r packages, include=TRUE}
require(ggplot2)
library(ggplot2)
require(RCurl)
library(RCurl)
```

#pulling data
```{r data, include=TRUE}
url <- getURL("https://raw.githubusercontent.com/choudhury1023/CUNY-MSDA-Bridge-Workshop-R/master/birthwt.csv")
birthwt <- read.csv(text = url)
birthwt <- birthwt[,c(2, 6, 11)]
```

#rename columns
```{r rename, include=TRUE}
names(birthwt) <- c("under_weight", "smoker", "birth_weight")
head(birthwt)
```

#change values
```{r values, include=TRUE}
birthwt$smoker <- factor(c("No", "Yes")[birthwt$smoker + 1])
birthwt$under_weight <- factor(c("No", "Yes")[birthwt$under_weight + 1])
summary(birthwt)
```

#boxplot for birth weight
```{r boxbw, include=TRUE}
ggplot(birthwt, aes(1, birth_weight)) + geom_boxplot(aes(fill = factor(smoker)))
```

#histogrm for birth wight distribution
```{r histbw, include=TRUE}
ggplot(birthwt, aes(x=birth_weight, fill=smoker)) + geom_histogram(binwidth=100)
```

#add column to identify rows containing somker mother and under weight child
```{r addcolumn, include=TRUE}
birthwt$smokerunderwt <-  birthwt$smoker=="Yes" & birthwt$under_weight=="Yes"
birthwt
```

#filtering out rows non smoker
```{r smoker, include=TRUE}
smokerwt <- birthwt$smoker== "Yes"
S <- birthwt[smokerwt,]
S
summary (S)
```

#find smoker mean weight manually
```{r smean, include=TRUE}
smean <- mean(subset(birthwt$birth_weight, birthwt$smoker== "Yes"))
smean
```
#filtering out rows smoker
```{r nonsmoker, include=TRUE}
nonsmokerwt <- birthwt$smoker== "No" 
N <- birthwt[nonsmokerwt,]
N
summary (N)
```

#find non smoker mean weight manually
```{r nonsmean, include=TRUE}
nonsmean <- mean(subset(birthwt$birth_weight, birthwt$smoker=="No"))
nonsmean
```

#t.test
```{r ttest, include=TRUE}
t.test (birthwt$birth_weight[birthwt$smoker == "Yes"], 
        birthwt$birth_weight[birthwt$smoker == "No"])
#the t.test shows the diffrence of mean is 170.1 and a p-value of 0.007003 which is lowr than 0.05, we will reject null hypothesis and take alternative hypothesis. In conclusion we can say there is a relationshipe between mother smoking during pregnancy and baby being under weight at birth.
```

