---
title: "Aircraft Data"
author: "Ahsanul Choudhury"
date: "August 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Aircraft Data

#Description

#Aircraft Data, deals with 23 single-engine aircraft built over the years 1947-1979, from Office of Naval Research. The dependent variable is cost (in units of \$100,000) and the explanatory variables are aspect ratio, lift-to-drag ratio, weight of plane (in pounds) and maximal thrust.

#Source: http://vincentarelbundock.github.io/Rdatasets/datasets.html

#required packages
```{r packages, include=TRUE}
require(ggplot2)
library(ggplot2)
require(RCurl)
library(RCurl)
```

#pulling data
```{r data, include=TRUE}
url <- getURL("https://raw.githubusercontent.com/choudhury1023/CUNY-MSDA-Bridge-Workshop-R/master/aircraft.csv")
aircraft <- read.csv(text = url)
aircraft <- aircraft[,c(2:6)]
```
#rename columns
```{r rename, include=TRUE}
names(aircraft) <- c("aspect_ratio", "lift_to_drag_ratio", "weight", "maximal_thrust","cost")
head(aircraft)
```
#Generate summary level descriptive statistics
```{r summary, include=TRUE}
summary(aircraft)
```

#R graphics

#Histogram for Aspect Ratio
```{r histA, include=TRUE}
qplot(aspect_ratio, data = aircraft, geom="histogram", ,binwidth = 0.5,main = "Histogram for Aspect Ratio", xlab = "Aspect Ratio", fill=I("blue"), col=I("black"), xlim=c(1, 7))
```

#Histogram for Lift to Drag Ratio
```{r histL, include=TRUE}
qplot(lift_to_drag_ratio, data = aircraft, geom="histogram", ,binwidth = 0.5,main = "Histogram for Lift to Darg Ratio", xlab = "Lift to Drag Ratio", fill=I("green"), col=I("black"), xlim=c(1, 10))
```

#Boxplot for Cost
```{r boxC, include=TRUE}
ggplot(aircraft, aes(y = cost, x = 1)) + geom_boxplot(fill = "yellow")
```

#Scatter plot for Weight and Maximal Thrust
```{r scattwt, include=TRUE}
ggplot(aircraft, aes(weight, maximal_thrust))+ geom_point(aes(colour = I("red")))
#A relationship is visible, havier aircrafts are likely to have more thrust 
```